\herschel\footnote{{\it Herschel} is an ESA space observatory with science instruments provided by European-led Principal Investigator consortia and with important participation from NASA.} Orion Protostar Survey (HOPS) Imaging and Photometry Results.

### 2017-10-17 HOPS/photometry/ms.tex
final comments from Will
toms comments and changes

\section{Data Processing \& Map-making}
\label{sec:dp}
\par
We start data processing at the so-called level~1 stage from the Herschel Science Archive \citep{pacspipeline}.  Level~1 contains calibrated timelines\footnote{Timelines are simply readouts from individual pixels ordered sequentially by time of observation.  Note that for our observations the PACS instrument was used with a fixed readout frequency of 10 Hz.} from individual PACS bolometers from which all instrumental effects have already been removed except for the low frequency noise component (the so-called 1/f noise\footnote{The so-called 1/f noise modifies the signal timelines by adding a drift component whose amplitude is a power-law function of its Fourier frequency.}).  Our processing mitigates the 1/f noise, combines the two independent orthogonal scan directions, and projects the timelines onto the final image of the field.  We will refer to these steps as 'map-making' hereafter.  The processing steps leading up to level~1 are described in \citet{pacs} and in the data processing guides for the Herschel Interactive Processing System, HIPE \cite{hipe}.  All data discussed here are based on the FM7 version of the PACS calibration \citep{pacscal} and processed with version 9 of the HIPE software.  Our final maps have spatial scales of 1.6\arcsec/pixel and 3.2\arcsec/pixel for the 70~\micron\ and 160~\micron\ PACS filters, respectively.  We used two different approaches for map-making that are described below.

\subsection{The High Pass Filter (HPF) branch}
\par
This technique filters bolometer timelines and, as the name suggests, blocks all temporal frequencies lower than the filter width (conversely, it allows only temporal frequencies smaller than the filter width).  Such filtering removes the low-frequency signal present in the timelines from both the 1/f noise as well as from astrophysical sources.  The primary advantage is that point sources, whose temporal frequencies are higher than the chosen filter width, are preserved.  Thus, this branch is useful for point source photometry.  The HPF filtering was applied as follows:  First, for any given readout in the timeline, the median value is calculated within a window of preceding and following readouts.  Only those readouts that are not flagged as a glitch or otherwise identified as problematic are included in the calculation.  Second, this median value is subtracted from the signal value in the current readout.  The process is repeated for all readouts in the timeline.  We investigated several different HPF filter window widths and settled on 15 readouts (1.5 seconds) and 20 readouts (2 seconds) for the 70~\micron\ and 160~\micron\ filters, respectively.  These provided the optimal balance between preserving the signal from point sources and filtering out as much of the 1/f noise as possible.  We use the HPF branch as implemented in HIPE and described by \cite{hpf}.
\par
For the HPF filtering, we assume that the median value determines the local sky emission and any variations in this median value are purely due to the drift caused by 1/f noise.  However, there is substantial amount of spatially extended emission in our fields from the molecular cloud itself (this emission is also referred to as nebulosity).  This widespread emission has the undesired effect of altering the median values.  Thus, along with the 1/f noise, the extended spatial emission (nebulosity) is also removed by the HPF processing  because the value of the local sky (taken as the median, see above) includes emission from local nebulosity.  In addition, point sources themselves may also elevate the median value in the HPF filter.  Thus, we mask and exclude point sources from the median calculation.  These masks are generated from the first iteration of HPF map-making and used in the 2nd iteration.  We use maps only from the 2nd iteration for our analysis.  We investigate the effect of nebulosity on the HPF and its consequences on the subsequent photometry in Section~\ref{sec:phot}

@@

\section{Photometry}
\label{sec:phot}
\par
Photometry is notoriously difficult in star-forming regions in particular.  Specific challenges are: (i) differentiating spatially unresolved knots in the nebular emission from actual point sources, (ii) estimating the local background contribution.  The nebular emission is usually complex and contains gradients of emission at all spatial scales that violate background homogeneity assumptions commonly used in aperture photometry and Point Spread Function (PSF)-fitted photometry algorithms.  (iii) Disentangling close binaries.  Further, protostars have a higher binary frequency than their main-sequence counterparts \citep{binaries}.  We, therefore, rely on multiple approaches in both map-making and photometry to resolve these issues.  The final list of photometry values is determined by combining the multiple approaches by considering the relative merits (strengths and weaknesses) of each approach for each source individually.

\subsection{Aperture photometry}
\par
\cite{pbrpaper}\ and \cite{furlan} describe the details of the aperture photometry procedure we followed for our images.  For convenience, we briefly reprise the salient points here:  We used the HPF branch images for aperture photometry measurements.  To avoid nebular contamination from the local environment, we place the inner annulus of our apertures as spatially close to the source as possible.  This step necessitates customized aperture corrections since the background annuli include a significant fraction of a source's point spread function (PSF) profile.  We used the Vesta calibration image (PACS's PSF standard, see Lutz \etal\ 2012) to calculate aperture corrections.  We set the inner radius of sky annulus to the aperture radius to ensure the
sky annulus sample the spatially varying nebulosity near the source.  The adopted values for 70~\micron\ are 9.6\arcsec, and 19.2\arcsec, for the aperture and sky annuli, respectively. The aperture correction factor is 0.7331.  In the 160~\micron\ images, the adopted values are 12.8\arcsec, and 25.6\arcsec, for the aperture and sky annuli, respectively.  The aperture correction factor at 160~\micron\ is 0.6602.
\par
The aperture photometry technique has several factors in its favor: First, it is a reliable and well-understood technique.  Next, it is the technique used for flux calibration of the PACS instrument \citep{pacscal}.  Further, the use of narrow apertures as described above alleviates nebular contamination and source crowding issues for most sources.  However, the complex structure in the images means that brightnesses for a significant fraction of sources are not well measured using aperture photometry.  The primary issue is the presence of a source or strong nebular emission in either the aperture or the sky annulus.  The sky annulus can be particularly affected by the presence of a strong source given that we have opted to use fairly narrow sizes for them (hence, small number of pixels are available for sky estimation).   These issues affect the 160~\micron\ image more strongly because there is simply more emission detected at that wavelength.  For these subset of sources, it was necessary to use a secondary photometry technique that is less susceptible to the issues noted above.  The next two subsections describe our method for identifying contaminated aperture photometry sources.

@@

\section{Results}
\label{sec-results}
\par
In this contribution, we focus on the {\em observed}\ properties of the target protostars.  A series of companion publications, \cite{will}, \cite{furlan}, \cite{erin}, and \cite{pbrpaper} present results and analysis from other scientific investigations and interpretations using these data from the HOPS program.   We further restrict our attention to the observed differences between the sub-regions of both Orion~A and B clouds.  In particular, we have elected to examine the following relationships: (i) flux distributions, (ii) color distributions, and, (iii) slopes of the SEDs.  We augment our analysis with previously published results for the PBR sources \cite{pbrpaper}, and sub-mm photometry \citep{thomas}.  We present results from each of the above identified observation quantity below, and collectively discuss the picture of star-formation implied by these observations in Section~\ref{sec:discussion}.

@@

\par
\subsection{Flux distribution functions}
\label{lfs}
\par
Figure~\ref{fig:fd}\ shows the flux distribution (FD) functions at 70~\micron\ and 160~\micron, respectively, for all sources detected in our sample.  Figures~\ref{fig:fd70regional}~\&~\ref{fig:fd160regional} show the flux distributions segregated by sub-region, and separately for the Orion~A (left column) and Orion~B (right column) clouds.  The sub-region and the number of associated sources are identified within the sub-panels of the Figures.  We do not consider upper limits in these FDs (sources with flag value = 2 in Table~\ref{tbl:obs}). There are distinct differences apparent in the observed flux distributions as a function of sub-region.  We note that in a flux-limited survey, such as ours, local environmental differences in the amount and structure of the nebular emission can significantly bias source detection and photometry algorithms.  We quantify FD differences and investigate the role of bias below.
\par
The 70~\micron\ FDs are relatively flat (\ie, the number of sources in all bins is similar) for the Orion~B sub-regions: LDN~1622, NGC~2068, and NGC~2023/4.  This situation is likely due to the low numbers of protostars available for each sub-region.  In Orion~A, the 70~\micron\ FDs have an apparent peak at or near 1 Jy in all sub-regions.  And, no significant differences are obvious between sub-regions, except for LDN~1641-N.
\par
The 160~\micron\ FDs show clear and obvious differences amongst the sub-regions.  The following results, in particular, are noted: (i) there is a complete lack of sources below $\sim$500~mJy in OMC-2/3 region.  (ii) The LDN~1641 (all) and OMC~2/3 sub-regions have apparent peaks in FDs at different flux values.  (iii) The peak of the ONC-S flux distribution is between those of LDN~1641 and OMC~2/3.  And, (iv) in Orion~B, LDN~1622 and NGC~2023/4 show differences in the observed distribution, even with the low number of sources available.
\par
The 160~\micron\ FDs suggest a trend of increasingly bright sources towards lower declinations in Orion~B, and increasingly bright sources towards higher declinations in Orion~A.  Since the sub-regions are segregated by declination in both the Orion~A and B clouds, declination is a useful proxy for sub-region.
We investigate this apparent trend by plotting the median flux for each sub region as a function of declination.  Figure~\ref{fig:mdFlux_v_dec} shows the resulting quantitative correlations.  The median shows a systematic change as a function of declination (sub-region) as noted above.  The trend is, however, weakly noted in the 70~\micron\ fluxes, and shows OMC 2/3 region with significantly different median flux for the 160~\micron\ fluxes.  The implications for this result are discussed in Section~\ref{sec:discussion}.

@@

\acknowledgments
\par
BA was partially supported with NASA grant IPAC.ALI-OTKP-1-JPL.000094 for this work.  The work of W. F. was supported in part by an appointment to the NASA Postdoctoral Program at Goddard Space Flight Center, administered by Oak Ridge Associated Universities through a contract with NASA.
\par
The work of AMS was partially supported by the Deutsche Forschungsgemeinschaft priority program 1573 ("Physics of the Interstellar Medium").
\par
J. Tobin acknowledges support provided by NASA through Hubble Fellowship grant HST-HF-51300.01-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS 5-26555. The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc.
\par
The Herschel spacecraft was designed, built, tested, and launched under a contract to ESA managed by the Herschel/Planck Project team by an industrial consortium under the overall responsibility of the prime contractor Thales Alenia Space (Cannes), and including Astrium (Friedrichshafen) responsible for the payload module and for system testing at spacecraft level, Thales Alenia Space (Turin) responsible for the service module, and Astrium (Toulouse) responsible for the telescope, with in excess of a hundred subcontractors.
\par
PACS has been developed by a consortium of institutes led by MPE (Germany) and including UVIE (Austria); KU Leuven, CSL, IMEC (Belgium); CEA, LAM (France); MPIA (Germany); INAF-IFSI/OAA/OAP/OAT, LENS, SISSA (Italy); IAC (Spain). This development has been supported by the funding agencies BMVIT (Austria), ESA-PRODEX (Belgium), CEA/CNES (France), DLR (Germany), ASI/INAF (Italy), and CICYT/MCYT (Spain).
\par
HIPE is a joint development by the Herschel Science Ground Segment Consortium, consisting of ESA, the NASA Herschel Science Center, and the HIFI, PACS and SPIRE consortia.

>>>>

\section{Data Processing \& Map-making}
\label{sec:dp}
\par
We start data processing at the level~1 stage from the Herschel Science Archive \citep{pacspipeline}.  Level~1 contains calibrated timelines\footnote{Timelines are simply readouts from individual pixels ordered sequentially by time of observation.  Note that for our observations the PACS instrument was used with a fixed readout frequency of 10 Hz.} from individual PACS bolometers from which all instrumental effects have already been removed except for the low frequency noise component (the so-called 1/f noise\footnote{The so-called 1/f noise modifies the signal timelines by adding a drift component whose amplitude is a power-law function of its Fourier frequency.}).  Our subsequent processing mitigates the 1/f noise, combines the two independent orthogonal scan directions for each OBSID group, and projects the timelines onto the final image of the field.  We will refer to these steps as 'map-making' hereafter.  The processing steps leading up to level~1 are described in \citet{pacs} and in the data processing guides for the Herschel Interactive Processing System, HIPE \cite{hipe}.  All data discussed here are based on the FM7 version of the PACS calibration \citep{pacscal} and processed with version 9 of the HIPE software.  Our final maps have spatial scales of 1.6\arcsec/pixel and 3.2\arcsec/pixel for the 70~\micron\ and 160~\micron\ PACS filters, respectively.  We used two different approaches for map-making that are described below.

\subsection{The High Pass Filter (HPF) branch}
\par
This technique filters bolometer timelines and, as the name suggests, blocks all temporal frequencies lower than the filter width (conversely, it allows only temporal frequencies smaller than the filter width).  Such filtering removes the low-frequency signal present in the timelines from both the 1/f noise as well as from astrophysical sources.  The primary advantage is that point sources, whose temporal frequencies are higher than the chosen filter width, are preserved.  Thus, this branch is useful for point source photometry.  The HPF filtering is applied as follows:  First, for any given readout in the timeline, the median value is calculated within a user-selected window of preceding and following readouts.  Only those readouts that are not flagged as a glitch or otherwise identified as problematic are included in the calculation.  Second, this median value is subtracted from the signal value in the current readout.  The process is repeated for all readouts in the timeline.  We investigated several different HPF filter window widths and settled on 15 readouts (1.5 seconds) and 20 readouts (2 seconds) for the 70~\micron\ and 160~\micron\ filters, respectively.  These provided the optimal balance between preserving the signal from point sources and filtering out as much of the 1/f noise as possible.  We use the HPF branch as implemented in HIPE and described by \cite{hpf}.
\par
For the HPF filtering, we assume that the median value determines the local sky emission and any variations in this median value are purely due to the drift caused by 1/f noise.  However, there is substantial amount of spatially extended emission in our fields from the molecular cloud itself (this emission is also referred to as nebulosity).  This widespread emission has the undesired effect of altering the median values.  Thus, along with the 1/f noise, the extended spatial emission (nebulosity) is also removed by the HPF processing  because the value of the local sky (taken as the median, see above) includes emission from local nebulosity.  In addition, point sources themselves may also elevate the median value in the HPF filter.  In our map-making, we mask and exclude point sources from the median calculation.  These masks are generated from the first iteration of HPF map-making and used in the 2nd iteration.  We use maps only from the 2nd iteration for our analysis.  We investigate the effect of nebulosity on the HPF and its consequences on the subsequent photometry in Section~\ref{sec:phot}

@@

\section{Photometry}
\label{sec:phot}
\par
Photometry is notoriously difficult in star-forming regions.  Specific challenges are: (i) differentiating spatially unresolved knots in the nebular emission from actual point sources, (ii) estimating the local background contribution.  The nebular emission is usually complex and contains gradients of emission at all spatial scales that violate background homogeneity assumptions commonly used in aperture photometry and Point Spread Function (PSF)-fitted photometry algorithms.  (iii) Disentangling close binaries.  Further, protostars have a higher binary frequency than their main-sequence counterparts \citep{binaries}.  We, therefore, rely on multiple approaches in both map-making and photometry to resolve these issues.  The final list of photometry values is determined by combining the multiple approaches by considering the relative merits (strengths and weaknesses) of each approach for each source individually.

\subsection{Aperture photometry}
\par
\cite{pbrpaper}\ and \cite{furlan} describe the details of the aperture photometry procedure we followed for our images.  For convenience, we briefly reprise the salient points here:  We used the HPF branch images for aperture photometry measurements.  To avoid nebular contamination from the local environment, we place the inner annulus of our apertures as spatially close to the source as possible.  This step necessitates customized aperture corrections since the background annuli include a significant fraction of a source's point spread function (PSF) profile.  We used the Vesta calibration image (PACS's PSF standard, see Lutz \etal\ 2012) to calculate aperture corrections.  We set the inner radius of the sky annulus to the aperture radius to ensure the
sky annulus samples the spatially varying nebulosity near the source.  The adopted values for 70~\micron\ are 9.6\arcsec, and 19.2\arcsec, for the aperture and sky outer annulus radii, respectively. The aperture correction factor is 0.7331.  In the 160~\micron\ images, the adopted values are 12.8\arcsec, and 25.6\arcsec, for the aperture and sky annuli, respectively.  The aperture correction factor at 160~\micron\ is 0.6602.
\par
The aperture photometry technique has several factors in its favor: First, it is a reliable and well-understood technique.  Next, it is the technique used for flux calibration of the PACS instrument \citep{pacscal}.  Further, the use of narrow apertures as described above reduces the effect of nebular contamination and source crowding issues for most sources.  However, the complex structure in the images means that brightnesses for a significant fraction of sources are not well measured using aperture photometry.  The primary issue is the presence of flux from a neighboring source or strong extended emission in either the aperture or the sky annulus.  The sky annulus can be particularly affected by the presence of a strong source given that we have opted to use fairly narrow sizes for them (hence, small number of pixels are available for sky estimation).   These issues affect the 160~\micron\ image more strongly because there is more extended emission detected at that wavelength.  For these subset of sources, it was necessary to use a secondary photometry technique that is less susceptible to the issues noted above.  The next two subsections describe our method for obtaining photometry in crowded or confused fields.

@@

\section{Results}
\label{sec-results}
\par
In this contribution, we focus on the {\em observed}\ photometric properties of the target protostars, focusing on the fluxes in four bands: the 24~\micron\ from \spitzer/MIPS and the 0, 100, and 160~\micron\ bands from \herschel/PACS instruments..  A series of companion publications, \cite{will}, \cite{furlan}, \cite{erin}, and \cite{pbrpaper} present results and analysis from other scientific investigations and interpretations using these data from the HOPS program.   Our analysis concentrates on the spatial variations in the flux and color distributions across the Orion clouds.  To search for variations in the color and flux distributions, we divide the Orion clouds into X distinct sub-regions, shown in Table~\ref{tbl:regions}.  We augment our analysis with previously published results for the PBR sources \cite{pbrpaper}, and sub-mm photometry \citep{thomas}.  We present results from each of the above identified observation quantity below, and collectively discuss the implications for star-formation implied by these observations in Section~\ref{sec:discussion}.

@@

\par
\subsection{Flux distribution functions}
\label{lfs}
\par
Figure~\ref{fig:fd}\ shows the flux distribution (FD) functions at 70~\micron\ and 160~\micron, respectively, for all sources detected in our sample.  Figures~\ref{fig:fd70regional}~\&~\ref{fig:fd160regional} show the flux distributions segregated by sub-region, and separately for the Orion~A (left column) and Orion~B (right column) clouds.  The sub-region and the number of associated sources are identified within the sub-panels of the Figures.  We do not consider upper limits in these FDs (sources with flag value = 2 in Table~\ref{tbl:obs}). There are distinct differences apparent in the observed flux distributions as a function of sub-region.  We note that in a flux-limited survey, such as ours, local environmental differences in the amount and structure of the nebular emission can significantly bias source detection and photometry algorithms.  We quantify FD differences and investigate the role of bias below.
\par
The 160~\micron\ FDs show clear and obvious differences amongst the sub-regions.  The following results, in particular, are noted: (i) there is a complete lack of sources below $\sim$500~mJy in OMC-2/3 region.  (ii) The LDN~1641 (all) and OMC~2/3 sub-regions have apparent peaks in FDs at different flux values.  (iii) The peak of the ONC-S flux distribution is between those of LDN~1641 and OMC~2/3.  And, (iv) in Orion~B, LDN~1622 and NGC~2023/4 show differences in the observed distribution, even with the low number of sources available.
\par
The 160~\micron\ FDs suggest a trend of increasingly bright sources towards lower declinations in Orion~B, and increasingly bright sources towards higher declinations in Orion~A.  Since the sub-regions are segregated by declination in both the Orion~A and B clouds, declination is a useful proxy for sub-region.
We investigate this apparent trend by plotting the median flux for each sub region as a function of declination.  Figure~\ref{fig:mdFlux_v_dec} shows the resulting quantitative correlations.  The median shows a systematic change as a function of declination (sub-region) as noted above.  The trend is, however, weakly noted in the 70~\micron\ fluxes, and shows OMC 2/3 region with significantly different median flux for the 160~\micron\ fluxes.  The implications for this result are discussed in Section~\ref{sec:discussion}.

@@

\acknowledgments
\par
BA was partially supported with NASA grant IPAC.ALI-OTKP-1-JPL.000094 for this work.
\par
The work of AMS was partially supported by the Deutsche Forschungsgemeinschaft priority program 1573 ("Physics of the Interstellar Medium").
\par
J. Tobin acknowledges support provided by NASA through Hubble Fellowship grant HST-HF-51300.01-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS 5-26555. The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc.
\par
The Herschel spacecraft was designed, built, tested, and launched under a contract to ESA managed by the Herschel/Planck Project team by an industrial consortium under the overall responsibility of the prime contractor Thales Alenia Space (Cannes), and including Astrium (Friedrichshafen) responsible for the payload module and for system testing at spacecraft level, Thales Alenia Space (Turin) responsible for the service module, and Astrium (Toulouse) responsible for the telescope, with in excess of a hundred subcontractors.
\par
PACS has been developed by a consortium of institutes led by MPE (Germany) and including UVIE (Austria); KU Leuven, CSL, IMEC (Belgium); CEA, LAM (France); MPIA (Germany); INAF-IFSI/OAA/OAP/OAT, LENS, SISSA (Italy); IAC (Spain). This development has been supported by the funding agencies BMVIT (Austria), ESA-PRODEX (Belgium), CEA/CNES (France), DLR (Germany), ASI/INAF (Italy), and CICYT/MCYT (Spain).
\par
HIPE is a joint development by the Herschel Science Ground Segment Consortium, consisting of ESA, the NASA Herschel Science Center, and the HIFI, PACS and SPIRE consortia.
